Synthesizing Symmetric Lenses

### 2018-07-09 papers/SymmetricLenses/main.tex
add termination condition descritpion improvement

In contrast to the identity lens, the \Disconnect lens does not propagate edits
from one format to the other. The disconnect lens takes four arguments: two
regular expressions and two strings. The regular expressions specify the formats
of the corresponding data on the two sides while the strings provide default
values. On creates, the input values are thrown away, and default values are
returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
example, the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them\bcp{Potentially confusing to
  readers (and, even to me, hacky). Why not relate each of them to the constant
  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
  like UNK used before in practice, but if it makes the example cleaner and more
  comprehensible, sure.  Anybody else have thoughts on the matter?}\bcp{No,
  I meant why not concatenate two disconnect lenses, one relating salary to
  the empty string and one relating the empty string to company.}.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

We now provide the typing derivations and semantics for these lenses. For
simplicity of presentation,\bcp{Is there any known issue with relaxing this
  restriction? I've often found ti convenient to use id lenses with ambiguous
  REs...}\afm{no known issues, I just didn't want to go into specifics about
  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
semantics only over unambiguous regular expressions, even though some lenses can
be defined over ambiguous regular expressions (for example, the identity lens
doesn't need any ambiguity constraints).

@@

The procedure \PCF{\Continue} terminates when the sum of the number of regular expressions of the
smallest unproposed size with the distance of those regular expressions from the
originals exceeds the cost of the current best lens. At this point, \RXSearch
stops proposing regular expression pairs, and instead returns to the user the
best lens found thus far. Note this means that if \RXSearch finds a bijective
lens, which has zero cost it will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs, other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

@@

We included a number of optimizations in our algorithm. We added ``open'' and
``closed'' annotations to regular expressions, an expansion inference algorithm,
hash-consing and memoization, and compositional synthesis to our algorithm. We
also enable an additional parameter in our specifications that customizes how
permissive \RXSearch{} is in accepting proposed lenses.

@@

\paragraph*{\RXSearch Permissiveness Customization}
In this work, we are constrained by three competing constraints. (1) We want to
return lenses quickly to users, (2) we want to return good lenses, and (3) we
don't want to require users to write a large number of examples in the
specifications. Our termination heuristic works well for most situations, but
there are situations where a large number of expansions must be performed to
find the correct lens.

@@

While our defaults termination condition quickly finds the correct lens for
\afm{35} of our \afm{45} benchmarks, our termination condition is not
one-size-fits all. Particularly complex lenses, or lenses that project a
particularly large amount of data, can be synthesized more easily and more
quickly by customizing how permissive \RXSearch is in returning lenses.
% end synthesis

>>>>

In contrast to the identity lens, the \Disconnect lens does not propagate edits
from one format to the other. The disconnect lens takes four arguments: two
regular expressions and two strings. The regular expressions specify the formats
of the corresponding data on the two sides while the strings provide default
values. On creates, the input values are thrown away, and default values are
returned ($\CreateROf{\String} =
\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
example, the \lstinline{salary} and \lstinline{company} fields are independent,
so we use the disconnect lens to relate them\bcp{Potentially confusing to
  readers (and, even to me, hacky). Why not relate each of them to the constant
  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
  like UNK used before in practice, but if it makes the example cleaner and more
  comprehensible, sure.  Anybody else have thoughts on the matter?}.
The defaults are strings that indicate the information is missing.
%
\begin{lstlisting}
disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
\end{lstlisting}
%
The insert \lstinline{ins} and delete \lstinline{del} lenses are
syntactic sugar for uses of the disconnect lens in which a string
constant is omitted entirely from the source and target formats,
respectively.
\begin{lstlisting}
 ins($t$) = disconnect("", $t$, "", $t$)
 del($s$) = disconnect($s$, "", $s$, "")
\end{lstlisting}

@@

We now provide the typing derivations and semantics for these lenses. For
simplicity of presentation,\bcp{Is there any known issue with relaxing this
  restriction? I've often found ti convenient to use id lenses with ambiguous
  REs...}\afm{no known issues, I just didn't want to go into specifics about
  ambiguity in an already-full paper}, we define the typing derivations and
semantics only over unambiguous regular expressions, even though some lenses can
be defined over ambiguous regular expressions (for example, the identity lens
doesn't need any ambiguity constraints).

@@

The procedure \PCF{\Continue} returns false based on the how long the search has
been going, and how hard it expects the next class of problems to be. In
particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
current best lens has cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c >= d + \Log_2(n)
\end{gather*}

When $\Call{Continue}{pq}$ returns $\False$, \RXSearch stops proposing regular
expression pairs, and instead returns to the user the best lens found thus far.
Note this means that if \RXSearch finds a bijective lens, which has zero cost it
will immediately return.
%
\subsection{Stochastic DNF Regular Expressions}
Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
for stochastic regular expressions.  Of all the star-semiring equations
required to relate two \SDNFREabbrevs, only those relating to Kleene star
(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
two \SDNFREabbrevs, other axioms, like associativity, commutativity
or distributivity, are unneeded.  This important property makes it possible to
construct an efficient algorithm ($\GreedySynth$), directed
by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
same ``class'' (\IE, any lenses that have the same type when the
Kleene star axioms are unused).  This property, and the resulting
efficiency of $\GreedySynth$ is key to the effectiveness of our
core algorithms.

@@

We included a number of optimizations in our algorithm. We added ``open'' and
``closed'' annotations to regular expressions, an expansion inference algorithm,
hash-consing and memoization, and compositional synthesis to our algorithm. We
also enable an additional parameter in our specifications that customizes the
termination condition for our procedure.

@@

\paragraph*{Termination Customization}
In this work, we are constrained by three competing constraints. (1) We want to
return lenses quickly to users, (2) we want to return good lenses, and (3) we
don't want to require users to write a large number of examples in the
specifications. Our termination heuristic works well for most situations, but
there are situations where a large number of expansions must be performed to
find the correct lens.

@@

For these types of situations, users can input a custom number, $t$, to
customize the strictness of the termination condition. In particular, if
$\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at distance $d$, the
number of pairs in $pq$ at distance $d$ is $n$, and the current best lens has
cost $c$:
\begin{gather*}
  \Call{Continue}{pq} := c >= \Min(d + \Log_2(n) - t, 0)
\end{gather*}
With a positive $t$, the termination condition is more strict, and the algorithm
will spend longer searching for more likely lenses. With a negative $t$ the
termination condition is less strict, and the algorithm will spend less time
searching for more likely lenses.  Our default $t$ is set to $0$.

While our defaults termination condition quickly finds the correct lens for
\afm{35} of our \afm{45} benchmarks, our termination condition is not
one-size-fits all. Particularly complex lenses, or lenses that project a
particularly large amount of data, can be synthesized more easily and quickly by
customizing how permissive \RXSearch is in returning lenses.
% end synthesis
